---
title: "Your Title"
author: "Your Name"
format: 
  html:
    embed-resources: true
---

## Section 2.2: Basic Plot Structure

A ggplot plot requires three things:

1. the name of the data set (data = _____).

2. an aesthetic mapping: maps variables in the data set to components of the plot.

3. a geom (gives the plot type).

```{r}
library(tidyverse)
stat113_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/stat113.csv")
```

## Section 2.3: Introduction to Plotting

Single Quant. Variable:

```{r}
ggplot(data = stat113_df, aes(x = GPA)) +
  geom_histogram()

ggplot(data = stat113_df, aes(x = GPA)) +
  geom_freqpoly(bins = 17)

ggplot(data = stat113_df, aes(x = GPA)) +
    geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 17) +
    geom_freqpoly(bins = 17, colour = "skyblue3")
```

Single Cat. Variable

```{r}
ggplot(data = stat113_df, aes(x = Award)) +
  geom_bar(colour = "tomato", fill = "coral")
```

Two Categorical Variables

```{r}
ggplot(data = stat113_df, aes(x = Year, fill = SocialMedia)) +
  geom_bar() 

ggplot(data = stat113_df, aes(x = Sex, fill = Sport)) +
  geom_bar(position = "fill") +
  labs(y = "Proportion")
```

One cat., one quant.

```{r}
ggplot(data = stat113_df, aes(x = Award, y = Exercise)) +
  geom_boxplot()

ggplot(data = stat113_df, aes(x = Award, y = Exercise)) +
  geom_violin()
## see more shape

ggplot(data = stat113_df, aes(x = Award, y = Exercise)) +
  geom_jitter(width = 0.10)
## rough idea of sample size

ggplot(data = stat113_df, aes(x = Sport, y = GPA)) +
  geom_violin(colour = "red", fill = "orange")
```

## Section 2.4: Practice

Class Exercise 1.

```{r}
ggplot(data = stat113_df, aes(x = SocialMedia)) +
  geom_bar()

stat113_restructured <- stat113_df |> group_by(SocialMedia) |>
  summarise(n_social = n())
stat113_restructured

ggplot(data = stat113_restructured, aes(x = SocialMedia)) +
  geom_bar()
## still counting up the number of rows each level 
## of SocialMedia appears (it's now just 1 for each level)

ggplot(data = stat113_restructured, aes(x = SocialMedia, y = n_social)) +
  geom_col()
## geom_col(): useful for bar plots when the counts are explicitly
## given in another column.
```

Class Exercise 2

```{r}
fitness_full <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv",
                         col_types = list(stepgoal = col_factor())) |>
  mutate(weekend_ind = case_when(weekday == "Sat" | weekday == "Sun" ~ "weekend",
                                 .default = "weekday"))
fitness_full
ggplot(data = fitness_full, aes(x = distance, y = active_cals)) +
  geom_point()
```

Class Exercise 3.

```{r}
ggplot(data = fitness_full, aes(x = Date, y = steps)) +
  geom_line() +
  geom_smooth()
```

Class Exercise 4.

```{r}
ggplot(data = stat113_df, aes(x = Exercise, y = GPA)) +
  geom_point() +
  facet_wrap(~ Sex)
## no, the relationship doesn't change based on Sex

ggplot(data = stat113_df, aes(x = Exercise, y = GPA, colour = Sex)) +
  geom_point() 
## same conclusion: no, the relationship does not change much based on Sex.

## combo:
ggplot(data = stat113_df, aes(x = Exercise, y = GPA, colour = Sex, shape = Sport)) +
  geom_point()

## another option
ggplot(data = stat113_df, aes(x = Exercise, y = GPA)) +
  geom_point() +
  facet_grid(Sex ~ Sport)
## facet_grid can grid a couple of categorical variables to form the facets
```

Class Exercise 5.

```{r}
ggplot(data = stat113_df, aes(x = Award, y = Exercise)) +
  geom_boxplot(outlier.shape = 8)
```

Your Turn 1.

a. 20 rows (one per student) and at least one column (for the major).

b. 3 rows (one for each level/category of major) and 2 columns (one column for the levels and one column for the counts).

Your Turn 2.

```{r}
ggplot(data = stat113_df, aes(x = GPA, y = TV)) +
  geom_point(aes(colour = Year))

ggplot(data = stat113_df, aes(x = GPA, y = TV)) +
  geom_point() +
  facet_wrap(~ Year)

## for this example, the faceted version looks better/cleaner.
## more generally, the more levels that a categorical variable has, the worse colour becomes (harder to distinguish and plots look messier).

ggplot(data = stat113_df, aes(x = Wgt, y = Hgt)) +
  geom_point() +
  facet_wrap(~ Sex)
ggplot(data = stat113_df, aes(x = Wgt, y = Hgt)) +
  geom_point(aes(colour = Sex)) 
```

Your Turn 3.

```{r}
fitness_full

## a
ggplot(data = fitness_full, aes(x = distance)) +
  geom_freqpoly(aes(colour = month))

## b
ggplot(data = fitness_full, aes(x = distance)) +
  geom_histogram() +
  facet_wrap(~ month)

## c
ggplot(data = fitness_full, aes(x = month, y = distance)) +
  geom_boxplot()
## for this exaqmple, boxplots look best, but it sure
## would be nice to put the months in chronological order!
```

Your Turn 5.

```{r}
ggplot(data = stat113_df, aes(x = GPA, y = Exercise)) +
  geom_line()

## the line plot worked for the fitness full data set because there was exactly one y value for each unique x, and, the x's were app;roximately equally spaced. This happens the most when x is time!
## this isn't the case for Exercise and GPA so a scatterplot is much better!
```

## Section 2.6: More Practice

```{r}
stat113_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/stat113.csv")
glimpse(stat113_df)
## glimpse() is especially nice for data sets with a 
## lot of columns
```

Class Exercise 1.

```{r}
ggplot(data = stat113_df, aes(x = Wgt, y = Hgt,
                              colour = Sport)) +
  geom_point() +
  geom_smooth()

ggplot(data = stat113_df, aes(x = Wgt, y = Hgt)) +
  geom_point(aes(colour = Sport)) +
  geom_smooth()
## moving colour = Sport to the geom_point() function
## makes the colour aesthetic local, not global.
## So, the points use the colour aesthetic but the 
## smoother does not.

```

Class Exercise 2.

```{r}
ggplot(data = stat113_df, aes(x = TV)) +
  geom_histogram(colour = "black", fill = "white")
```

Your Turn 1.

a. No: method = "lm" should not be in aes()

b. Correct.

c. No: colour = "purple" should not be in aes().

d. No: colour = groupvar needs to be in aes() since groupvar is a variable in the data set.

e. Correct.

f. No: colour = "purple" should not be in aes().

Your Turn 2.

Specify part (e) as only using local aesthetics (outside of a code chunk).

e.
 all aesthetics supplied locally:
 
ggplot(data = toy_df) +
  geom_point(aes(x = xvar, y = yvar), colour = "purple") +
  geom_smooth(method = "lm", aes(x = xvar, y =yvar, colour = groupvar))