---
title: "Chapter 3: Wrangling with dplyr"
author: "Your Name"
format: 
  html:
    embed-resources: true
---

## Section 3.1

## Section 3.2: Practice

```{r}
library(tidyverse)
library(babynames)
babynames
```

Class Exercise 1 and 2.

```{r}
babynames_m10 <- babynames |> filter(year == 2017 & sex == "M") |>
  arrange(desc(n)) |>
  slice(1:10)
## 1:10 is 1, 2, 3, 4, 5, 6, 7, 8, 9, 10

## class exercise 2.
ggplot(data = babynames_m10, aes(x = name, y = n)) +
  geom_col()
```

Class Exercises 3 and 4.

```{r}
babynames_matthew <- babynames |> filter(name == "Matthew" & sex == "M")
ggplot(data = babynames_matthew, aes(x = year, y = n)) +
  geom_line()
```

Your Turn 1 and Your Turn 2.

Your Turn 1.

```{r}
babynames |> filter(year != 1945)

babynames |> filter(n >= 400 & n <= 5000)


```

Your Turn 2.

```{r}
babynames_mh <- babynames |> filter((name == "Monica" | name == "Hilary") & sex == "F")
ggplot(data = babynames_mh, aes(x = year, y = n)) +
  geom_line(aes(colour = name))
```

Your Turn 3.

```{r}
babynames_5 <- babynames |> filter((name == "Matthew" & sex == "M") |
                      (name == "Jessica" & sex == "F") |
                      (name == "Ivan" & sex == "M") |
                      (name == "Robin" & sex == "M") |
                      (name == "Lillian" & sex == "F"))
ggplot(data = babynames_5, aes(x = year, y = n)) +
  geom_line(aes(colour = name))
```

Your Turn 4.

```{r}
babynames_12 <- babynames |> filter(year == 1991 & sex == "M") |>
  arrange(desc(n)) |>
  slice(1:12) ## keep 12 most popular

ggplot(data = babynames_12, aes(x = name, y = n)) +
  geom_col(colour = "seagreen4", fill = "seagreen1")
```

## Section 3.4

Class Exercise 1.

```{r}
slumajors_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/SLU_Majors_17_21.csv")

slumajors_df <- slumajors_df |>
  mutate(ntotal = nfemales + nmales) |>
  mutate(percfemale = 100 * nfemales / (nfemales + nmales))

slumajors_df <- slumajors_df |>
  mutate(morewomen = if_else(percfemale > 50,
                             true = "Yes",
                             false = "No")) |>
  mutate(large_majority =
           case_when(percfemale >= 70 ~ "female",
                     percfemale <= 30 ~ "male",
                     percfemale > 30 & percfemale < 70 ~ "none")) 

slumajors_df

slumajors_df |> mutate(testcase = case_when(percfemale >= 40 ~ "Yes",
                                            percfemale > 60 ~ "No"))
## if a row satisfies more than one condition, then case_when() uses
## the __first__ condition that's satisfied to determine the output

```

Class Exercise 2.

```{r}
## example of rank
slumajors_df |> mutate(rank_percfemale = rank(percfemale)) |>
  select(rank_percfemale, everything())
## smallest value of percfemale gets a rank of 1,
## second smallest value of percfemale gets a rank of 2, etc.

slumajors_df |> mutate(rank_percfemale = rank(desc(percfemale))) |>
  select(rank_percfemale, everything())
## ranking percfemale so that the largest value gets rank of 1

slumajors_df |> group_by(morewomen) |>
  mutate(rank_percfemale = rank(percfemale)) |>
  select(rank_percfemale, everything())
## now ranking within each group established by the "morewomen" variable
```

```{r}
babynames |> group_by(year, sex) |>
  mutate(rank_pop = rank(desc(n))) |>
  filter(rank_pop == 1) |>
  ungroup() |> ## removes the year and sex grouping
  group_by(name) |>
  summarise(n_most_pop = n())  |> ## number of rows for each name
  arrange(desc(n_most_pop))
```

Class Exercise 3.

```{r}
babynames |> group_by(year, sex) |>
  filter(n == max(n)) |>
 ## mutate(rank_pop = rank(desc(n))) |>
 ## filter(rank_pop == 1) |>
  ungroup() |> ## removes the year and sex grouping
  group_by(name) |>
  summarise(n_most_pop = n())  |> ## number of rows for each name
  arrange(desc(n_most_pop))

babynames |> group_by(year, sex) |>
  arrange(desc(n)) |>
  slice(1) |>
 ## mutate(rank_pop = rank(desc(n))) |>
 ## filter(rank_pop == 1) |>
  ungroup() |> ## removes the year and sex grouping
  group_by(name) |>
  summarise(n_most_pop = n())  |> ## number of rows for each name
  arrange(desc(n_most_pop))
```

Class Exercise 4.

```{r}
babynames_test <- babynames |>
  group_by(year, sex) |> mutate(ntest = n / prop)
babynames_test |> slice(1, 2, 3, 4, 5)

## the issue is that we still have groups established (year and sex)
## so we are actually getting the first 5 rows for each group
276 * 5
```

Class Exercise 5.

```{r}
fitness_df <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv") |>
  select(-stepgoal)
fitness_df
```

```{r}
fitness_df |> mutate(step_10 = if_else(steps >= 10000,
                                       true = 1, 
                                       false = 0)) |>
  summarise(prop_10 = mean(step_10)) |>
  select(prop_10, everything())
## mean of a variable that is only 1's and 0's is the proportion of 1's.

## b.

fitness_df |> mutate(step_10 = if_else(steps >= 10000,
                                       true = 1, 
                                       false = 0)) |>
  group_by(year) |>
  summarise(prop_10 = mean(step_10)) |>
  select(prop_10, everything())


```

Your Turn 2.

```{r}
toy_df <- tibble(x = c(NA, 3, 4, 7),
                 y = c(1, 4, 3, 2),
                 z = c("A", "A", "B", NA))
toy_df

## a
toy_df |> mutate(test_x = x +5)
## rows with NAs for a variable in a mutate will give an NA.

## b
toy_df |> arrange(x)
toy_df |> arrange(desc(x))
## NAs go at the end.

## c
toy_df |> filter(x > 5)
## NAs get removed as part of this filter.

## d
toy_df |> summarise(median_x = median(x))
## returns a missing value since at least one x is missing

## if you want to drop the missing value
toy_df |> filter(!is.na(x)) |>
  summarise(median_x = median(x))
## OR
toy_df |> summarise(median_x = median(x, na.rm = TRUE))

## E
toy_df |> group_by(z) |>
  summarise(median_x = median(x, na.rm = TRUE))
## missing values become their own group
```

Your Turn 3. 

a and b. Generally safe: we are not restructuring the data set, removing important information. we are just creating a new variable (mutate) or sorting the rows (arrange).

c. not safe to do in general (removing data and overwriting the name of the original data frame means that you never have access to the data you removed again).

d. not safe to do in general: you have access to the restructured summary but not the original data set anymore.

e. similar to filter, not safe in general (unless you were absolutely sure you never needed access to those columns again).

Your Turn 4.

```{r}
slumajors_df |>
  summarise(meantotalmajor = mean(ntotal),
            totalgrad = sum(ntotal)) 
slumajors_df |>
  mutate(meantotalmajor = mean(ntotal),
            totalgrad = sum(ntotal)) |>
  select(meantotalmajor, totalgrad, everything())

## summarise() only keeps the variables specified in the summarise() statement
## as well as any grouping variables whereas mutate() keeps all variables.

## summarise() collapses this data to only keep a single row (since there is no grouping) whereas mutate() always preserves all rows in the data set.
```
