---
title: "Chapter 4: Communication"
author: "Your Name"
execute:
  warning: false
format: 
  html:
    fig-height: 4
    embed-resources: true
---

## Section 4.4: Practice

```{r}
#| echo: false
library(gapminder)
library(tidyverse)
country_df <- gapminder
```

Class Exercise 1.

Hint: use a group aesthetic in your ggplot geom.

```{r}
## figuring out which countries we wnt to label
country_df |> filter(continent == "Europe" & lifeExp < 50)
## Turkey

country_df |> filter(continent == "Africa" & lifeExp < 30)
## Rwanda

country_df |> filter(continent == "Asia" & year == 1977 & lifeExp < 40)
## Cambodia

## dplyr to make a data set that has rows for the three countries

country_3 <- country_df |> filter(country == "Cambodia" | country == "Rwanda" | country == "Turkey")

## next task: make a data set from country_3 that only has three rows (for our three labels)

country_3_min <- country_3 |> group_by(country) |>
  arrange(lifeExp) |>
  slice(1)
## OR
country_3 |> group_by(country) |>
  filter(lifeExp == min(lifeExp))

ggplot(data = country_df, aes(x = year, y = lifeExp)) +
  geom_line(aes(group = country)) +
  facet_wrap(~ continent) +
  geom_text(data = country_3_min, aes(label = country))

## task: figure out what countries correspond to the three
## "unusual" lines and to then directly label the plot
## with the country's name on the lines.

## instead of labeling the countries directly, try using colour
## to distinguish these three countries of interest:
## Hint: need to use a local data set as part of a second geom_line()
ggplot(data = country_df, aes(x = year, y = lifeExp)) +
  geom_line(aes(group = country), alpha = 0.2) +
  facet_wrap(~ continent) +
  geom_line(data = country_3, aes(colour = country),
            linewidth = 1.5) +
  scale_colour_viridis_d() +
  theme_minimal()

## alpha controls transparancy.
## can use colour to emphasize a few interesting data points (countries).


```

Class Exercise 8. Done above.

## Your Turn 

Your Turn 1.

Outliers were removed and he did not record that he removed them.

Using t.test() without writing down that that is the function that he used (and any arguments).

changing the labels and limits using point and click is also not reproducible.

Your Turn 2.

Hint:

```{r}
#| eval: false
## can add a column to the data set with mutate() where you 
## just specify the values yourself
## e.g., if df had just two rows:
df |> mutate(label_var = c("label 1", "label 2"))
```

```{r}
fitness_full <- read_csv("https://raw.githubusercontent.com/highamm/ds234_quarto/main/data_online/higham_fitness_clean.csv",
                         col_types = list(stepgoal = col_factor())) |>
  mutate(weekend_ind = case_when(weekday == "Sat" | weekday == "Sun" ~ "weekend",
                                 .default = "weekday"))

fitness_label <- fitness_full |> filter(Date == "2019-04-14" | 
                         Date == "2021-05-24") |>
  mutate(event = c("half-marathon", "grand canyon"))
ggplot(data = fitness_full,
       aes(x = Date, y = steps)) +
  geom_line() +
  geom_text(data = fitness_label, aes(label = event))
```

Your Turn 3.

