---
title: "Chapter 5: Workflow"
author: "Your Name"
execute:
  warning: false
format: 
  html:
    fig-height: 4
    embed-resources: true
---

## Section 5.7

Class Exericse 1. Because you're not recording the packages that you've used in your script window (so that someone else looking at your analysis doesn't know what packages you've used).

Class Exercise 2.

```{r}
library(tidyverse)
library(palmerpenguins)

ggplot(data = mtcars, aes(x = hp, y = drat)) +
  geom_point(aes(colour = factor(gear))) +
  facet_wrap(~ cyl) +
  geom_smooth()
```

Class Exercise 3.

```{r}
library(palmerpenguins)
penguins |> mutate(flipper_ratio = flipper_length_mm / body_mass_g) |>
  group_by(species, island) |>
  summarise(mean_flipper = mean(flipper_ratio, na.rm = TRUE)) |>
  arrange(mean_flipper) |>
  pivot_wider(names_from = c("species"), values_from = "mean_flipper")
```

Class Exercise 4.

```{r}
library(here)
videogame_df <- read_csv(here("data/videogame_clean.csv"))

## outliers are kept in
videogame_df |> summarise(mean_pt = mean(median_playtime, na.rm = TRUE),
                          median_pt = median(median_playtime, na.rm = TRUE))

## leave unfinished
```

Class Exercise 5.

```{r}
library(tidyverse)
library(here)
evals_df <- read_csv(here("data/evals_prof_S21.csv"))
evals_df
```

```{r}
## how can I split the Term variable into a 
## semester column and a year column?

library(tidyverse)
toy_reprex <- tibble(Term = c("F19", "F20", "S20", "S20"))
toy_reprex
```


Your Turn 3.

```{r}
library(palmerpenguins)
penguins |> mutate(flipper_ratio = flipper_length_mm / body_mass_g) |>
  relocate(flipper_ratio) |>
  filter(flipper_ratio > median(flipper_ratio, na.rm = TRUE)) |>
  group_by(species) |>
  summarise(count_var = n())
```


Your Turn 4.

a. don't know about filter()

```{r}
library(tidyverse)

reprex_df <- tibble(x = c(1, 5, 3, 6, 7, NA, 10),
                    y = c(4, 1, 2, NA, 4, 7, 8))
reprex_df

## how can I remove the rows that have any missing values
## for any variable in the data set?
```

Notes about here() and working directories.

R project establishes that the working directory starts in the same folder as the R project.

here() works with an R project. here() says when we read or write a data set to start where the R project is.

for example,

read_csv(here("data/alcohol.csv"))

says to start looking for the data folder "here" at the r project location.