---
title: "Chapter 6: `tidyr`"
execute:
  warning: false
format: 
  html:
    fig-height: 4
    embed-resources: true
---

```{r}
library(tidyverse)
library(here)
```

Your Turn 1.

```{r}
baseball_df <- read_csv(here("data/mlb2016.csv"))
baseball_df
```

First, the salary, total value, and avg annual columns all contain units ($).

The `Years` column contains really three different variables: the contract duration, the start year, and the end year (for tidy data, each column should contain just one variable, not three).

Your Turn 2 and Your Turn 3 (fix the Years column so that it's tidy).

```{r}
baseball_separated <- baseball_df |> separate(Years, c("duration", "years"),
                        sep = " ") |>
  separate(years, c("start", "end"), sep = "-")
## class exercise 3: some years were single contract years without a dash
## to separate by, which is what the wraning says.
```

Your Turn 4.

```{r}
baseball_tidy <- baseball_separated |>
  mutate(Salary = parse_number(Salary),
         Total.Value = parse_number(Total.Value),
         Avg.Annual = parse_number(Avg.Annual),
         Start = parse_number(start),
         End = parse_number(end))
baseball_tidy |> view()
## parse_number(x) says to only keep numbers in the column x and drop
## everything else (letters, symbols, anything that's not a number)
```

Your Turn 5.

```{r}
ggplot(data = baseball_tidy, aes(x = POS, y = Salary)) +
  geom_boxplot(colour = "skyblue4", fill = "skyblue1") +
  theme_minimal()

## with salary having dollar signs
##ggplot(data = baseball_df, aes(x = POS, y = Salary)) +
##  geom_boxplot(colour = "skyblue4", fill = "skyblue1") +
##  theme_minimal()

## in general, bar plots are not good to make when we have a categorical
## and a quantitative variable. They are useful for counts.
ggplot(data = baseball_tidy, aes(x = POS, y = Salary)) +
  geom_col()

ggplot(data = baseball_tidy, aes(x = POS, y = Salary)) +
  geom_jitter(width = 0.05, aes(colour = POS), show.legend = FALSE) +
  scale_colour_viridis_d() +
  theme_minimal()
## let's you see sample size!
## show.legend = FALSE removes the colour legend: in this case, the colour is redundant with the x-axis.
```

Your Turn 6.

```{r}
nfl_df <- read_csv(here::here("data/nfl_salary.csv"))
nfl_df
```

variables are position and salary. Each of these two variables should be in their own column (not spread out across a bunch of different columns).

Your Turn 7.

```{r}
nfl_df |> pivot_longer(c(Cornerback, `Defensive Lineman`,
                         Linebacker, `Offensive Lineman`))

nfl_tidy <- nfl_df |> pivot_longer(2:11, names_to = "position",
                                   values_to = "salary")
nfl_tidy
```

Your Turn 8.

```{r}
nfl_max <- nfl_tidy |> group_by(position, year) |>
  summarise(max_salary = max(salary))

nfl_tidy |> group_by(position, year) |>
  arrange(desc(salary)) |>
  slice(1)

nfl_tidy |> group_by(position, year) |>
  filter(salary == max(salary))

ggplot(data = nfl_max, aes(x = year, y = max_salary)) +
  geom_line(aes(colour = position))
## too many colours

ggplot(data = nfl_max, aes(x = year, y = max_salary)) +
  geom_line() +
  facet_wrap(~ position)

```